Continuous estimates of heat emission at Mt. Ruapehu using the Unscented Kalman Smoother

### 2022-09-30 manuscript/manuscript.tex
adding extra sections at the end

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass[doublespacing, linenumbers]{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
%\usepackage{graphicx}

@@

\def\includegraphic{}
\def\includegraphics{}

@@

\section*{Acknowledgements}
The authors would like to thank Nico Fournier for providing feedback and ideas
throughout the project and Sophie Pearson-Grant for her constructive and
insightful feedback on this manuscript. GNS Science's volcano monitoring group
has been an invaluable team to test and improve this work. Graphs have been
produced using the open-source graphing libraries Plotly
(https://plotly.com/python/), matplotlib (https://matplotlib.org/) and cartopy
(https://scitools.org.uk/cartopy/docs/latest/).

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Author's contributions}
  Y.B. wrote the main manuscript text, produced graphs, and implemented the
  Kalman Smoothing. S.S. worked on data collection and interpretation. T.H.
  developed and implemented the mass and energy balance model. All
  authors reviewed the manuscript.

\section*{Availability of data and materials}%% if any
The source code for this work can be accessed at
https://github.com/yannikbehr/pumahu. All data used in this study is provided
free of charge by the GeoNet program (https://www.geonet.org.nz)

@@

\begin{figure}[h!]
    %\includegraphics[width=12 cm]{figures/overview.png}  
    \caption{Image of Mt. Ruapehu's crater lake from 2008 (\textcopyright Visual
        Media Library, GNS Science, ID 6277, photographer Karen Britten). The red
        circle marks the location of the lake outlet. The location of Mt. Ruapehu is
        shown as a red triangle on the inset map.}
      \label{overview}
\end{figure}

\begin{figure}[h!]
   %\includegraphics[width=12 cm]{figures/sensitivity_analysis.png}  
    \caption{Surface energy loss ($Q_e$) with respect to lake water temperature
             ($T$). The model parameters that have been held fixed are shown in  
             the title.}
      \label{sens}
\end{figure}

\begin{figure}[h!]
	%\includegraphics[width=12 cm]{figures/synthetic_inversion_uks.png}  
    \caption{Inversion result for the synthetic test described in Section
        \ref{syn_test}. Turquoise lines and markers represent synthetic
        observations; red dashed lines are the inversion results and black
        dashed lines are the true input to the synthetic test. Error bars and
        shaded areas represent $\pm$ 3 standard deviations.}
  \label{syn_example}
\end{figure}

\begin{figure}[h!]
 	%\includegraphics[width=12 cm]{figures/time_frequency_analysis.png}  
    \caption{ a) Inference results for the heat input rate $\dot{Q_i}$ at RCL
             between 4 March 2016 and 1 February 2022. The solid line shows the
             marginal probability and the uncertainty is displayed as shaded
             region; b) Continuous wavelet transform of the time series shown in
             a; c) Power spectral density of the heat input rate shown as a
             solid line with uncertainty displayed as shaded region.}
\label{inference_result}
\end{figure}

\begin{figure}[h!]
 	%\includegraphics[width=12 cm]{figures/cumulative_heat.png}  
    \caption{Cumulative difference to the mean for the heat input rate into RCL
             (solid line) and its uncertainty (shaded region) calculated from
             the time-series in Figure \ref{inference_result}a.} 
\label{cumulative}
\end{figure}

>>>>

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
%\documentclass[doublespacing, linenumbers]{bmcart}
\documentclass[doublespacing]{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{graphicx}

@@

%\def\includegraphic{}
%\def\includegraphics{}

@@

\section*{Declarations}
\subsection*{Acknowledgements}
The authors would like to thank Nico Fournier for providing feedback and ideas
throughout the project and Sophie Pearson-Grant for her constructive and
insightful feedback on this manuscript. GNS Science's volcano monitoring group
has been an invaluable team to test and improve this work. Graphs have been
produced using the open-source graphing libraries Plotly
(https://plotly.com/python/), matplotlib (https://matplotlib.org/) and cartopy
(https://scitools.org.uk/cartopy/docs/latest/).

\subsection*{Funding}%% if any
This project was funded in parts by the Strategic Science Investment Funding to
GNS Science within the Datascience programme and by the GeoNet programme.

\subsection*{Abbreviations}%% if any
  \begin{tabular}{ll}
    RCL  & Ruapehu Crater Lake \\ 
    $\dot{M}$ & Lake water mass change \\
    $\dot{M_i}$ & Steam input \\ 
    $\dot{M_s}$ & Melt water and precipitation inflow \\
    $\dot{M_o}$ & Seepage and overflow \\ 
    $H$ & Enthalpy \\
    $T$ & Lake temperature \\
    $c_w$ & Specific heat of water \\
    $\dot{Q_i}$ & Energy input from volcanic sources \\
    $\dot{Q_r}$ & Solar irradiation \\
    $\dot{Q_e}$ & Surface energy loss \\
    $X$ & Total ion amount in the lake\\
  \end{tabular}

\subsection*{Availability of data and materials}%% if any
The source code for this work can be accessed at
https://github.com/yannikbehr/pumahu. All data used in this study is provided
free of charge by the GeoNet program (https://www.geonet.org.nz).

\subsection*{Ethics approval and consent to participate}%% if any
Not applicable

\subsection*{Competing interests}
The authors declare that they have no competing interests.

\subsection*{Consent for publication}%% if any
Not applicable

\subsection*{Author's contributions}
  Y.B. wrote the main manuscript text, produced graphs, and implemented the
  Kalman Smoothing. S.S. worked on data collection and interpretation. T.H.
  developed and implemented the mass and energy balance model. All
  authors reviewed the manuscript.

\subsection*{Authors' information}%% if any
All authors are affiliated with GNS Science. TH is Emeritus Volcano
Geophysicist, SS is Senior Volcano Geophysicist and YB is Volcano Science
Operations Specialist.

@@

\begin{figure}[h!]
    \includegraphics[width=12 cm]{figures/overview.png}  
    \caption{Image of Mt. Ruapehu's crater lake from 2008 (\textcopyright Visual
        Media Library, GNS Science, ID 6277, photographer Karen Britten). The red
        circle marks the location of the lake outlet. The location of Mt. Ruapehu is
        shown as a red triangle on the inset map.}
      \label{overview}
\end{figure}

\begin{figure}[h!]
   \includegraphics[width=12 cm]{figures/sensitivity_analysis.png}  
    \caption{Surface energy loss ($Q_e$) with respect to lake water temperature
             ($T$). The model parameters that have been held fixed are shown in  
             the title.}
      \label{sens}
\end{figure}

\begin{figure}[h!]
	\includegraphics[width=12 cm]{figures/synthetic_inversion_uks.png}  
    \caption{Inversion result for the synthetic test described in Section
        \ref{syn_test}. Turquoise lines and markers represent synthetic
        observations; red dashed lines are the inversion results and black
        dashed lines are the true input to the synthetic test. Error bars and
        shaded areas represent $\pm$ 3 standard deviations.}
  \label{syn_example}
\end{figure}

\begin{figure}[h!]
 	\includegraphics[width=12 cm]{figures/time_frequency_analysis.png}  
    \caption{ a) Inference results for the heat input rate $\dot{Q_i}$ at RCL
             between 4 March 2016 and 1 February 2022. The solid line shows the
             marginal probability and the uncertainty is displayed as shaded
             region; b) Continuous wavelet transform of the time series shown in
             a; c) Power spectral density of the heat input rate shown as a
             solid line with uncertainty displayed as shaded region.}
\label{inference_result}
\end{figure}

\begin{figure}[h!]
 	\includegraphics[width=12 cm]{figures/cumulative_heat.png}  
    \caption{Cumulative difference to the mean for the heat input rate into RCL
             (solid line) and its uncertainty (shaded region) calculated from
             the time-series in Figure \ref{inference_result}a.} 
\label{cumulative}
\end{figure}
